---
title: "MAPC ACCESSIBILITY-BASED TRAVEL MODEL"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, 
                      fig.pos = 'H', fig.cap=T)
library(tidyverse)
library(sf)
library(tmap)
library(shiny)
library(rpgcolorsr)
tmap_mode("view")
```

The [Metropolitan Area Planning Council (MAPC)](https://www.mapc.org/) has 
developed an accessibility-based sketch planning model for quick-hit analysis of
travel behaviors, including mode choice, non-motorized travel, average trip
length, trip distribution, and ride-hailing service utilization throughout the
Boston region. The sketch model responds to the need for reliable planning-level
insight into urban travel behaviors that are sensitive to land use, urban design,
and travel costs with minimal reliance on complex regional travel models. It 
utilizes **multimodal accessibility** as the overarching analysis framework, 
linking travel behavior to the effectiveness with which different travel modes 
(walking, driving, public transit, e.g.) connect households, jobs, schools, and 
shopping.

The sketch accessibility model was developed and piloted as part of MAPC's
analysis of the [West Station Area Transit Study](https://www.mapc.org/resource-library/west-station-area-transit-study/).
The map below shows the West Station study area. A [basic walkthrough](./west_station_exec.html)
of the West Station area existing conditions analysis, demonstrating
accessibility's relationship to travel behavior is available [here](./west_station_exec.html).
A summary of findings for several alternative future scenarios for the West
Station Area is available [here](.west_station_exec.html).


```{r, eval = TRUE}
zones <- read_sf("./data/mapping/window_blocks_simp.shp") %>% 
  filter(INFOCUS==1 & Base==1)
# Orange polys
    tm_shape(zones) + 
    tm_polygons(col=rpg_cols("orange"), alpha=0.5) +
# White borders
      tm_shape(zones) +
    tm_borders(col="#FFFFFF", lwd=2) +
    tm_basemap(leaflet::providers$Esri.WorldImagery)

```


## What is multimodal accessibility?

## {.tabset}

### Analysis components

Multimodal accessibility is an established transportation analysis metric that
is recently generating new insights into travel behavior. It is a relatively
simple metric, comprised of a few key components:

 - **Zones** are geographic aggregations, such as census blocks or travel
 analysis zones (TAZ's), that generalize locations where trips come from or go 
 to.
 
 - **Activities** are the reasons we travel: jobs, schools, stores, etc.
 
 - **Population groups** are defined by demographic characteristics that impact
 how, why, when, and where people travel.
 
 - **Costs** or **Impedances** refer to the time and money required to travel 
 from each zone to each other zone. Cost information is stored in **skims**,
 which are origin-destination matrices recording different cost components for
 quick summarization to total costs and/or application in cost-sensitive
 travel analyses.
 
 - **Decay** refers to the concept that travel is less likely to occur between
 two zones if the cost to travel between them is high. Specific rates of decay
 vary by mode and travel purpose.
 
 These components feature prominently in the [summary analysis of existing
 conditions in the West Station Area](./travel.html).

### Basic processing steps

The basic processing steps in developing multimodal accessibility scores 
include:

  - Estimate travel costs among zones. Costs may include travel time, fares,
  tolls, fuel costs, parking charges, and other expenses. 
  
  - Calculate generalized cost among zones as a function of each mode's relevant
  costs.
  
  - Apply decay formulas to estimate the "weight" of each OD pair, a rough
  approximation of travel probability based on generalized travel costs.
  
  - Broadcast destination-end activities into the decay matrix and multiply by
  decay weights to get each OD pair's "weighted activity" estimate.
  
  - Summarize weighted activity by origin zone to determine the access score.
  
The last two steps can be run on alternative axes so that access from origin-end
activities to each destination zone is summarized rather than access to 
destination-end activities from each origin zone.

These processing steps are codified in the [enhanced multimodal accessibility (emma)](https://renaissanceplanning.github.io/emma-docs/)
python library for generalized application and used in the 
[West Station analysis](./sphinx/html/index.html) python library developed for
MAPC's pilot study.


### Relationship to travel behavior

Multimodal accessibility offers an intuitive lens to understand the relative
impacts of planning alternatives on travel behavior. Access scores are
sensitive to:

  - local land use changes;
  - transportation costs and system performance (reflected in travel times);
  - policy variables (parking charges, e.g.). 
  
Access scores, in turn, can be used to assess **mode choice** and 
**trip distribution** by mode. In general, higher accessibility scores by a
given mode (especially relative to other modes) indicate higher likelihoods
that the mode will be used in daily travel.

Details of the sketch travel model developed for the Boston region using
accessibility are provided in the [Model Documentation](./travel.html). The
model estimates *trip generation*, *mode choice*, and *trip distribution* to 
assess travel impacts in a localized study area under alternative land use
and transportation scenarios.
