---
title: "MODE CHOICE MODEL SENSITIVITIES"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, 
                      fig.pos = 'H', fig.cap=T)
```

```{css}
caption {
 color: rgb(63, 78, 86);
 line-height: 1.5em;
 font-size: 16px;
 font-weight: bold;
}

p.caption {
 color: rgb(63, 78, 86);
 line-height: 1.5em;
 font-size: 16px;
 font-weight: bold;
}
```

```{r import}
library(knitr)
library(kableExtra)
library(readxl)
library(dplyr)
library(stringr)
```

This page highlights the factors that influence accessibility scores and mode
choice model outputs. The roster of key variables includes general indications
of their impact on model results. For more details on mode choice model
variables, see the [mode choice model report](./Mode_Choice_Report.html)

Four tables are presented in the tabs below:

 - **Access scores** are sensitive to the locations of jobs and households
 as well as the travel networks for each analyzed mode. This table focuses
 on the network attributes that influence accessibility scores. Since these
 scores are used in the mode choice models, these factors have a downstream
 impact on estimated mode shares.
 - **Production end mode choice** models combine access scores with household
 characteristics to estimate mode choice for the HBW, HBO, and HBSch purposes.
 - **Attraction end mode choice** models focus primarily on accessibility scores,
 with some consideration of the characteristics of households that can access
 a given zone alongside built environment variables.
 - **Non-home-based mode choice** models rely on access scores and the outputs
 of the attraction end mode choice models. When analyzing NHB trips, the mode of
 travel used to reach the non-home location is an important determinant of the
 likely mode of travel.

# Model component details {.tabset}

## Access scores
```{r access}
scores = readxl::read_xlsx("./data/Model_Sensitivities.xlsx",
               sheet = "Scores")
scores = data.frame(scores)
# scores = scores %>%
#   setNames(c("Factor","Nest","Influence")) %>%
#   mutate(Factor = str_replace_all(Factor, "\\u200B", ""),
#          Nest = str_replace_all(Nest, "\\u200B", ""),
#          Influence = str_replace_all(Influence, "\\u200B", ""))

kable(scores,
      booktabs = TRUE,
      longtable = TRUE,
      row.names = FALSE,
      col.names = c("Mode", "Cost Component", "Determinant of Cost", "Influence"),
      escape = FALSE,
      align = rep('l',times=3),
      caption = "Network attributes and cost parameters impacting accessibility scores") %>%
  kable_styling(bootstrap_options = c("striped", "hover",
                                      "responsive", "bordered")) %>%
  row_spec(0, bold=TRUE) %>% 
  collapse_rows(columns=1, valign="middle")
```

## Production end mode choice

```{r prods}
# https://stackoverflow.com/questions/49143690/rmarkdown-kable-vertically-align-cells
# centerText <- function(text){
#   paste0("\\multirow{1}{*}[0pt]{", text, "}")
# }

pe_hb = readxl::read_xlsx("./data/Model_Sensitivities.xlsx",
                  sheet = "PE_HB")
pe_hb = data.frame(pe_hb)
pe_hb = pe_hb %>%
  setNames(c("Factor","Nest","Influence")) %>%
  mutate(Factor = str_replace_all(Factor, "\\u200B", ""),
         Nest = str_replace_all(Nest, "\\u200B", ""),
         Influence = str_replace_all(Influence, "\\u200B", ""))

kable(pe_hb,
      booktabs = TRUE,
      row.names = FALSE,
      align = rep('l',times=3),
      caption = "Mode choice model sensitivities for production end home-based models") %>%
  kable_styling(bootstrap_options = c("striped", "hover",
                                      "responsive", "bordered")) %>%

  row_spec(0, bold=TRUE) %>% 
  collapse_rows(columns=1, valign="middle")
```

## Attraction end mode choice

```{r attrs}
ae = readxl::read_xlsx("./data/Model_Sensitivities.xlsx",
               sheet = "AE")
ae = data.frame(ae)
ae = ae %>%
  setNames(c("Factor","Nest","Influence")) %>%
  mutate(Factor = str_replace_all(Factor, "\\u200B", ""),
         Nest = str_replace_all(Nest, "\\u200B", ""),
         Influence = str_replace_all(Influence, "\\u200B", "")) %>%
  setNames(c("\\textbf{Factor}","\\textbf{Nest}","\\textbf{Influence}"))

kable(ae,
      booktabs = TRUE,
      longtable = TRUE,
      row.names = FALSE,
      col.names = c("Factor","Nest","Influence"),
      escape = FALSE,
      align = rep('l',times=3),
      caption = "Mode choice model sensitivities for attraction-end models") %>%
  kable_styling(bootstrap_options = c("striped", "hover",
                                      "responsive", "bordered")) %>%

  row_spec(0, bold=TRUE)  %>% 
  collapse_rows(columns=1, valign="middle")
```

## Non-home-based mode choice

```{r nhb}
pe_nhb = readxl::read_xlsx("./data/Model_Sensitivities.xlsx",
                   sheet = "PE_NHB")
pe_nhb = data.frame(pe_nhb)
pe_nhb = pe_nhb %>%
  setNames(c("Factor","Nest","Influence")) %>%
  mutate(Factor = str_replace_all(Factor, "\\u200B", ""),
         Nest = str_replace_all(Nest, "\\u200B", ""),
         Influence = str_replace_all(Influence, "\\u200B", ""))

kable(pe_nhb,
      booktabs = TRUE,
      row.names = FALSE,
      align = rep('l',times=3),
      caption = "Mode choice model sensitivities for production end non-home-based models") %>%
  kable_styling(bootstrap_options = c("striped", "hover",
                                      "responsive", "bordered")) %>%

  row_spec(0, bold=TRUE) %>% 
  collapse_rows(columns=1, valign="middle")
```




