---
title: "WEST STATION SCENARIO COMPARISONS"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, 
                      fig.pos = 'H', fig.cap=T)
library(tidyverse)
library(rpgcolorsr)
library(gridExtra)
library(stringr)
library(readxl)
library(scales)

# Pathing
setwd('../..')
root_path = getwd()
data_path = file.path(root_path, "FinalData")
# Inputs
input_path = file.path(data_path, "input", "zones")
lu_path = file.path(input_path, "lu")
# Outputs
tg_path = file.path(data_path, "lu")
net_path = file.path(data_path, "net")
scen_path = file.path(data_path, "scen")


```

This page provides summary comparisons of six different scenarios analyzed for
the West Station Area using the accessibility-based travel model. The summary
data and visualizations shown provide examples of the potential comparative
analyses supported by the model rather than an exhaustive menu of metrics.

The scenarios analyzed include:

1. Base - Approximates current land uses and travel networks

2. LRTP - Reflects adopted long term (2045) land use forecasts and cost feasible
travel networks.

3. RailVision - Reflects modified long term forecasts in the focus area and
expanded/enhanced regional rail services.

4. BRT-FEIR - Reflects modified long term forecasts in the focus area and 
potential bus rapid transit services.

5. BRT-Alt A - Reflects modified long term forecasts in the focus area and 
potential bus rapid transit services.

6. BRT-Alt B - Reflects modified long term forecasts in the focus area and 
potential bus rapid transit services.

Use the tabs below to compare scenario outputs across key model steps.

## {.tabset}
### Trip generation
In the trip generation step, households, jobs, school enrollments, etc. are used
to estimate how many trips begin or end in different zones. The charts below
present example summary outputs for four alternative land use configurations.
Base, LRTP, and FEIR are operative in the scenarios defined above. FEIR MAX is
a potential alternative land use forecast in the focus area used in early model
sensitivity testing, but it is not reflected in any reported scenarios.

The first chart shows trips productions and attractions generated in the window
area under each scenario, stratified by purpose. Generally, productions reflect
households and their trip-making trends based on various characteristics, like
income and vehicle ownership; attractions reflect jobs and school enrollments.
Both productions and attractions are forecasted to increase over base year 
conditions in all scenarios. Differences between the alternative forecasts are
subtle since they are isolated to the focus area, which is just a small subset
of the window area. The share of trips by each purpose is roughly consistent
accross all scenarios as well, with the most common travel purpose for productions
and attractions being home-based other trips.

```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(tg_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trip_gen_summary.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Purpose, End) %>% 
  summarize(Trips = sum(Trips))

trip_data$Purpose = factor(trip_data$Purpose,
                           levels=c("HBW", "HBO", "HBSch", "NHB"))
trip_data$Scenario = factor(trip_data$Scenario,
                            levels=c("Base", "LRTP", "FEIR", "FEIR_MAX"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=End, y=Trips, fill=Purpose), position="stack", stat="identity", color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(labels = c("Attractions", "Productions")) +
  labs(title = "Daily Trip Productions and Attractions in Window Area") + 
  facet_wrap(~Scenario, nrow = 1)
# Show the plot
trip_plot

```

Trip productions are sensitive to different household characteristics. The 
chart below shows the breakdown of households by vehicle ownership in the window
area in each alternative land use configuration. All future year scenarios show
a dramatic expected increase in one-car households throughout the window area.
Vehicle ownership is also an important factor in mode choice decisions. Having
high numbers of zero-car or one-car households will result in more non-auto
trips, all else being equal.

```{r, eval = TRUE}
# HH totals by vehicle ownership category
dir_names <- list.dirs(tg_path, recursive=FALSE)

# Create the data for plotting
hh_data = lapply(dir_names, function(x){
  fname = file.path(x, "hh_by_type_sum.csv")
  hh = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    mutate(HHSize= lapply(HHType, function(x){str_split(x, "_")[[1]][1]}) %>% unlist,
           Income= lapply(HHType, function(x){str_split(x, "_")[[1]][2]}) %>% unlist,
           Workers= lapply(HHType, function(x){str_split(x, "_")[[1]][3]}) %>% unlist,
           Vehicle= lapply(HHType, function(x){str_split(x, "_")[[1]][4]}) %>% unlist,
           Scenario = basename(x))
  return(hh)
}) %>% bind_rows

hh_sum <- hh_data %>%  group_by(Vehicle, Scenario) %>%  summarise(HH = sum(Households))
hh_sum$Scenario = factor(hh_sum$Scenario,
                         levels=c("Base", "LRTP", "FEIR", "FEIR_MAX"))

# Plot, with unique windows for each scenario 
hh_plot = ggplot(data = hh_sum) +
  geom_bar(
    aes(x=Scenario, y=HH, fill=Vehicle), position="stack", stat="identity", color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Households by Vehicle Ownership in Window Area")
hh_plot

```

### Accessibility scores
Accessibility scores describe the number of activities (jobs, school enrollments,
households, etc.) that can reach or be reached from a given zone. In addition
to informing mode choice models, the scores can be used to assess how well the
transportation system connects various activities to one another. The charts
below summarize average access to jobs scores for households by income group for
non-auto travel modes (transit, walking, and biking). The charts show that
households in the highest income bracket ($125,000 and higher) have the highest
access scores in the window area in all scenarios. Middle-income households 
($35,000 - $125,000) have the lowest non-auto access scores.

```{r, eval = TRUE}
# HH access by income
scens <- c("Base", "LRTP", "RailVision", "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B")
lu_configs <- c("Base", "LRTP", "FEIR", "FEIR", "FEIR", "FEIR")


# Create the data for plotting
access_data = map2(scens, lu_configs, function(scen_name, lu_name){
  # Read the excel file
  lu_file = file.path(input_path, lu_name, "MAPC_TAZ_data.xlsx")
  lu = readxl::read_xlsx(lu_file, "Zdata") %>% 
    select(TAZ, Income1, Income2, Income3, Income4)
  
  # Read the access file
  access_file = file.path(scen_path, scen_name, "access_to_jobs_transit_HBW.csv")
  access = read_csv(access_file) %>% 
    filter(INWINDOW == 1) %>% 
    select(TAZ, score = `Total Emp-Purpose-HBW`) %>% 
    mutate(Scenario = scen_name)
    
  # Join tables and summarize HH and weighted access scores
  sum_table <- access %>% 
    left_join(lu, by="TAZ") %>% 
    mutate(prod_Inc1 = Income1 * score,
           prod_Inc2 = Income2 * score,
           prod_Inc3 = Income3 * score,
           prod_Inc4 = Income4 * score) %>% 
    select(-c(TAZ, score)) %>% 
    group_by(Scenario) %>% 
    summarize_all(sum)
  
  # Get weighted average
  out_table <- sum_table %>% 
    mutate(`<$35k`  = prod_Inc1/Income1,
           `<$75k` = prod_Inc2/Income2,
           `<$125k` = prod_Inc3/Income3,
           `$125k+` = prod_Inc4/Income4) %>% 
    select(Scenario, `<$35k`, `<$75k`, `<$125k`, `$125k+`) %>% 
    gather("Income", "Score", 2:5)
  return(out_table)
}) %>% bind_rows

# Plot, with unique windows for each scenario 
access_data$Income = factor(access_data$Income, 
                           levels=c("<$35k", "<$75k","<$125k", "$125k+"))
access_data$Scenario = factor(access_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))
access_plot = ggplot(data = access_data) +
  geom_bar(
    aes(x=Income, y=Score, fill=Income), stat="identity", color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Average Transit Access to Jobs by Household Income") + 
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
access_plot

```



```{r, eval = TRUE}
# HH access by income
scens <- c("Base", "LRTP", "RailVision", "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B")
lu_configs <- c("Base", "LRTP", "FEIR", "FEIR", "FEIR", "FEIR")


# Create the data for plotting
access_data = map2(scens, lu_configs, function(scen_name, lu_name){
  # Read the excel file
  lu_file = file.path(input_path, lu_name, "MAPC_TAZ_data.xlsx")
  lu = readxl::read_xlsx(lu_file, "Zdata") %>% 
    select(TAZ, Income1, Income2, Income3, Income4)
  
  # Read the access file
  access_file = file.path(scen_path, scen_name, "access_to_jobs_walk_HBW_TAZ.csv")
  access = read_csv(access_file) %>% 
    filter(INWINDOW == 1) %>% 
    select(TAZ, score = `Total Emp-Purpose-HBW-Impedance-WalkTime`) %>% 
    mutate(Scenario = scen_name)
    
  # Join tables and summarize HH and weighted access scores
  sum_table <- access %>% 
    left_join(lu, by="TAZ") %>% 
    mutate(prod_Inc1 = Income1 * score,
           prod_Inc2 = Income2 * score,
           prod_Inc3 = Income3 * score,
           prod_Inc4 = Income4 * score) %>% 
    select(-c(TAZ, score)) %>% 
    group_by(Scenario) %>% 
    summarize_all(sum)
  
  # Get weighted average
  out_table <- sum_table %>% 
    mutate(`<$35k`  = prod_Inc1/Income1,
           `<$75k` = prod_Inc2/Income2,
           `<$125k` = prod_Inc3/Income3,
           `$125k+` = prod_Inc4/Income4) %>% 
    select(Scenario, `<$35k`, `<$75k`, `<$125k`, `$125k+`) %>% 
    gather("Income", "Score", 2:5)
  return(out_table)
}) %>% bind_rows

# Plot, with unique windows for each scenario 
access_data$Income = factor(access_data$Income, 
                           levels=c("<$35k", "<$75k","<$125k", "$125k+"))
access_data$Scenario = factor(access_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))
access_plot = ggplot(data = access_data) +
  geom_bar(
    aes(x=Income, y=Score, fill=Income), stat="identity", color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Average Walk Access to Jobs by Household Income") + 
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
access_plot

```


```{r, eval = TRUE}
# HH access by income
scens <- c("Base", "LRTP", "RailVision", "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B")
lu_configs <- c("Base", "LRTP", "FEIR", "FEIR", "FEIR", "FEIR")


# Create the data for plotting
access_data = map2(scens, lu_configs, function(scen_name, lu_name){
  # Read the excel file
  lu_file = file.path(input_path, lu_name, "MAPC_TAZ_data.xlsx")
  lu = readxl::read_xlsx(lu_file, "Zdata") %>% 
    select(TAZ, Income1, Income2, Income3, Income4)
  
  # Read the access file
  access_file = file.path(scen_path, scen_name, "access_to_jobs_bike_HBW_TAZ.csv")
  access = read_csv(access_file) %>% 
    filter(INWINDOW == 1) %>% 
    select(TAZ, score = `Total Emp-Purpose-HBW-Impedance-BikeTime`) %>% 
    mutate(Scenario = scen_name)
    
  # Join tables and summarize HH and weighted access scores
  sum_table <- access %>% 
    left_join(lu, by="TAZ") %>% 
    mutate(prod_Inc1 = Income1 * score,
           prod_Inc2 = Income2 * score,
           prod_Inc3 = Income3 * score,
           prod_Inc4 = Income4 * score) %>% 
    select(-c(TAZ, score)) %>% 
    group_by(Scenario) %>% 
    summarize_all(sum)
  
  # Get weighted average
  out_table <- sum_table %>% 
    mutate(`<$35k`  = prod_Inc1/Income1,
           `<$75k` = prod_Inc2/Income2,
           `<$125k` = prod_Inc3/Income3,
           `$125k+` = prod_Inc4/Income4) %>% 
    select(Scenario, `<$35k`, `<$75k`, `<$125k`, `$125k+`) %>% 
    gather("Income", "Score", 2:5)
  return(out_table)
}) %>% bind_rows

# Plot, with unique windows for each scenario 
access_data$Income = factor(access_data$Income, 
                           levels=c("<$35k", "<$75k","<$125k", "$125k+"))
access_data$Scenario = factor(access_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))
access_plot = ggplot(data = access_data) +
  geom_bar(
    aes(x=Income, y=Score, fill=Income), stat="identity", color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Average Bicycle Access to Jobs by Household Income") + 
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
access_plot

```


### Mode Choice
Mode choice estimates are a pivotal output of the sketch accessibility model.
Access scores and demographic data combine to estimate what proportions of 
trips are likely to be made by each travel mode. 

The chart below shows the breakdown of trips by mode and purpose for each
scenario for productions in the window area. Future year scenarios indicate
increased walking and biking trips, especially for non-home-based trips. Walk-
access transit (WAT) is much more common than drive-access transit (DAT), as
expected for the urban context of the window area. All scenarios indicate an
expected increase in WAT trips for the home-based work (HBW), but there is
substantial variety in transit utilization for home-based other (HBO) trips,
with shares highest in the RailVision scenario and lowest in the BRT-FEIR and
BRT-Scen B runs.


```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trips_by_mode_predist.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1, End == "P") %>%
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Purpose, Mode) %>% 
  summarize(Trips = sum(trips))

# Set factor columns for ordering plots
trip_data$Mode = factor(trip_data$Mode, 
                        levels=c("walk", "bike","WAT", "DAT", "passenger", "driver"))
trip_data$Purpose = factor(trip_data$Purpose,
                           levels=c("HBW", "HBO", "HBSch", "NHB"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=Purpose, y=Trips, fill=Mode), position="stack", stat="identity",
    color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "chartreuse", "orange", 
                                        "burnt_orange_ramp5", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Estimated Daily Trip Productions in Window Area by Mode") +
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```


The next chart shows transit ridership estimates by mode of access for 
productions and attractions separately. In all scenarios, there are more transit
attractions than productions, reflecting the large number of destinations in the
window area. These destinations can be reached by both WAT and DAT modes, whereas
DAT is less relevant to productions in the window area. Total expected transit
utilization is highest in the RailVision scenario.

```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trips_by_mode_predist.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    filter(Mode == "WAT" | Mode == "DAT") %>% 
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Mode, End) %>% 
  summarize(Trips = sum(trips))

# Set factor columns for ordering plots
trip_data$Mode = factor(trip_data$Mode, levels=c("DAT", "WAT"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=End, y=Trips, fill=Mode), position="stack", stat="identity",
    color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("burnt_orange_ramp5", "orange")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Estimated Daily Transit Trips To and From Window Area") +
  scale_x_discrete(labels = c("Attractions", "Productions")) +
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```

The chart below presents similar information for non-motorized trips (walking
and biking). All future scenarios anticipate fewer non-motorized attractions but
more non-motorized productions, reflecting increased residential development and
more travel to the window area by transit and auto modes from the remainder area.
The RailVision and BRT-Scen A scenarios have the greatest amount of non-motorized
trip-making.

```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trips_by_mode_predist.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    filter(Mode == "walk" | Mode == "bike") %>% 
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Mode, End) %>% 
  summarize(Trips = sum(trips))

# Set factor columns for ordering plots
trip_data$Mode = factor(trip_data$Mode, levels=c("bike", "walk"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=End, y=Trips, fill=Mode), position="stack", stat="identity",
    color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("chartreuse", "green")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Estimated Daily Non-Motorized Trips To and From Window Area") +
  scale_x_discrete(labels = c("Attractions", "Productions")) +
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```

### Distribution
The distribution phase estimates where trips are likely to travel to and from.
Outputs from this portion of the model can be summarized in terms of trip
lenghts, durations, and costs. The chart below compares vehicle miles of travel
(VMT) generated purpose across each scenario. This reflects the total number of
trips made by driving and the length of each trip. VMT is expected to increase
over the base year in all future year scenarios. This increase is more pronounced
for trips produced in the window area than for those attracted there. VMT 
estimates are highest in the BRT-FEIR and BRT-Scen B scenarios.

```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trip_len_dist_AM.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1 & Mode == "driver") %>%
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Purpose) %>% 
  summarize(Miles_From = sum(Miles_From, na.rm=TRUE),
            Miles_To = sum(Miles_To, na.rm=TRUE))

# Set factor columns for ordering plots
trip_data$Purpose = factor(trip_data$Purpose, 
                           levels=c("HBW", "HBO", "HBSch", "NHB"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))

trip_data = trip_data %>% gather("FromTo", "VMT", 3:4)

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=FromTo, y=VMT, fill=Purpose), position="stack", stat="identity",
    color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "orange", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Estimated Vehicle Miles of Travel To and From Window Area in AM Peak") +
  scale_x_discrete(labels = c("Trips To", "Trips From")) +
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```

The next chart displays average trip lengths (in miles) to and from the window
area by mode across all scenarios. In general, DAT trips are typically longer
than trips by other modes for travel to the window area. Walk trips tend to be
about a mile in length, while bike trips average about 3-4 miles. Trips attracted
to the window area tend to be longer than trips produced there. This reflects
the area's location near the urban core of the Boston region (allowing trips
produced there to be relatively short) and its attractions-heavy character
(which would indicate more trips from the remainder area coming to the window).
Overall, trip lengths appear to be longest in the BRT scenarios, though transit
trips to the window are longest in the RailVision scenario.

```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "trip_len_dist_AM.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Mode) %>% 
  summarize(Trips_From = sum(Trips_From, na.rm=TRUE),
            Miles_From = sum(Trip_Miles_From, na.rm=TRUE),
            Trips_To = sum(Trips_To, na.rm=TRUE),
            Miles_To = sum(Trip_Miles_To, na.rm=TRUE)
            ) %>% 
  mutate(Avg_From = Miles_From/Trips_From,
         Avg_To = Miles_To/Trips_To) %>%
  gather(key="FromTo", value="AvgTripLen", 7:8)


# Set factor columns for ordering plots
trip_data$Mode = factor(trip_data$Mode, 
                        levels=c("walk", "bike","WAT", "DAT", "passenger", "driver"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))
trip_data$FromTo = factor(trip_data$FromTo,
                          levels=c("Avg_To", "Avg_From"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=FromTo, y=AvgTripLen, fill=Mode), position="dodge", stat="identity",  color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "chartreuse", "orange", 
                                        "burnt_orange_ramp5", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Average Trip Length by Mode From Window Area in AM Peak") +
  scale_x_discrete(labels = c("Attractions", "Productions")) + 
  ylab("Miles") +
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```

### TNC
Finally, the TNC post-processing phase of the model can be summarized in terms
of TNC trip totals, stratified by purpose and/or mode replaced. The chart below
provides an example comparing TNC trip totals to and from the window area by the
mode being replaced. In all scenarios, transit trips are the most likely candidates
for TNC trip replacement. TNC estimates for all future year scenarios are higher
than the base year estimates, but there is substantial fluctuation. the LRTP and
BRT-Scen A scenarios suggest the highest TNC utilization, while the RailVision
and BRT-Scen B scenarios are only slightly higher than the base estimates.


```{r, eval = TRUE}
# Trips by Purpose and End for each scenario
dir_names <- list.dirs(scen_path, recursive=FALSE)

# Create the data for plotting
trip_data = lapply(dir_names, function(x){
  fname = file.path(x, "TNC_zone_summary_AM.csv")
  trips = read_csv(fname) %>%
    filter(INWINDOW == 1) %>%
    mutate(Scenario = basename(x))
  return(trips)
}) %>% bind_rows %>% 
  group_by(Scenario, Mode, FT) %>% 
  summarize(tnc_trips = sum(tnc_trips, na.rm=TRUE))

# Set factor columns for ordering plots
trip_data$Mode = factor(trip_data$Mode, 
                        levels=c("walk", "bike","WAT", "DAT", "passenger", "driver"))
trip_data$Scenario = factor(trip_data$Scenario,
                              levels=c("Base", "LRTP", "RailVision", 
                                       "BRT_FEIR", "BRT_Scen_A", "BRT_Scen_B"))

# Plot, with unique windows for each scenario 
trip_plot = ggplot(data = trip_data) +
  geom_bar(
    aes(x=FT, y=tnc_trips, fill=Mode), position="stack", stat="identity",  color="#ffffff") +
  scale_fill_manual(values = rpg_cols(c("green", "chartreuse", "orange", 
                                        "burnt_orange_ramp5", "blue", "shale")) %>% 
                         unname()) +
  scale_y_continuous(labels = comma) +
  labs(title = "TNC Trips To and From the Window Area by Mode Replaced") +
  xlab("TNC Trips") + 
  scale_x_discrete(labels = c("Trips To", "Trips From")) + 
  facet_wrap(~Scenario, nrow = 2)
# Show the plot
trip_plot
```





