---
title: "WEST STATION AREA ANALYSIS WALKTHROUGH"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, 
                      fig.pos = 'H', fig.cap=T)
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(shiny)
library(rpgcolorsr)
library(classInt)
library(mapdeck)
library(rosm, quietly=TRUE)

# Read shape data
map_path = getwd()
zones <- read_sf("./data/mapping/MAPC_Zones_simp.shp")
blocks <- read_sf("./data/mapping/window_blocks_simp.shp") %>% 
  filter(Base==1)

# Get zone tags
setwd('../..')
data_path = getwd()
zone_tags <- read_csv("./FinalData/input/model_area_zones.csv")
zones <- left_join(zone_tags, zones)

```

This page describes the key processing steps in the MAPC sketch 
accessibility-based travel model piloted in the West Station Area. The maps
presented provide examples of the modeling components to illustrate the logic
behind the model and indicate the kinds of outputs it can generate.

The West Station Area analysis focuses on four trip purposes: *home-based work*, 
*home-based school*, *home-based other*, and *non-home-based*; and six travel 
modes: *Walk*, *bike*, *auto driver*, *auto passenger*, *transit (walk access)*,
and *transit (drive access)*. This page presents a sampling of purpose and mode
results for illustrative purposes.

The sketch travel model is organized into seven steps. The steps are described
briefly below. Then maps showing the results of steps 3 through 7 are presented
for the West Station existing conditions analysis to demonstrate the workflow
and relationships among the steps.

1. **Skim cleaning** - csv files containing OD travel costs are
processed into a standard format for consistency across all scenarios.

2. **Import skims** - standardized OD skim csv files are imported into emma's
[Skim class](https://renaissanceplanning.github.io/emma-docs/od.html#skim).

3. **Summarize accessibility** - skim files, land use data, and decay curves
are used to calculate access to total jobs, retail jobs, and school enrollments
from origin zones and access from households by vehicle ownership and income
to destination zones.

4. **Trip generation** - land use data are used to estimate person trip
productions and attractions by TAZ. In the immediate vicinity around the 
West Station Area (referred to as the "Window" area), trips are disaggregated
from TAZ level to census block level.

5. **Mode choice** - Accessibility and demographic data are used to estimate 
the proportion of trips made by different modes. Mode shares are estimated
separately for trip productions and attractions. They are applied to trip
generation estimates to develop an initial estimate of trips by mode.

6. **Distribution** - Trip productions and attractions by mode are matched into 
OD pairs based on generalized OD costs. The resulting trip table is balanced
using iterative proportional fitting to ensure total trip productions and 
attractions in each zone match trip generation estimates.

7. **TNC trips** - Transportation Network Company (TNC) trips are estimated
based on distribution results, estimated TNC costs, and assumptions about which
trips are likely to switch to TNC.

## {.tabset}

### Summarize accessibility
In this step, access to and from various activities and population groups is
estimated for TAZ's and blocks (for walking and biking in the window area only).
Access scores inform mode choice models, while the decay curves used to estimate
access also inform trip distribution patterns.

The example maps below reflect access to all jobs by walking, biking, transit
(walk access), and car under existing conditions, using decay rates for the 
home-based work travel purpose.

```{r, eval = TRUE}
tmap_mode("view")
setwd(data_path)
# Read tables
zdata_a <- read_csv("./FinalData/Scen/Base/access_to_jobs_auto_HBW.csv")
zdata_t <- read_csv("./FinalData/Scen/Base/access_to_jobs_transit_HBW.csv")
zdata_w <- read_csv("./FinalData/Scen/Base/access_to_jobs_walk_HBW_TAZ.csv")
zdata_b <- read_csv("./FinalData/Scen/Base/access_to_jobs_bike_HBW_TAZ.csv")
bdata_w <- read_csv("./FinalData/Scen/Base/access_to_jobs_walk_HBW.csv", 
                    col_types =  cols(block_id = col_character()),
                    )
bdata_b <- read_csv("./FinalData/Scen/Base/access_to_jobs_bike_HBW.csv",
                    col_types =  cols(block_id = col_character())
                    )

# Join data
zones_a <- left_join(
  zdata_a, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_t <- left_join(
  zdata_t, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_w <- left_join(
  zdata_w, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_b <- left_join(
  zdata_b, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
blocks_w <- left_join(
  bdata_w, blocks, c("block_id" = "GEOID10", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
blocks_b <- left_join(
  bdata_b, blocks, c("block_id" = "GEOID10", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))

# Filter data
zones_w <- zones_w %>% filter(INWINDOW==0)
zones_b <- zones_b %>% filter(INWINDOW==0)

# Patch data
zones_w <- zones_w %>% rename(`Total Emp-Purpose-HBW` = `Total Emp-Purpose-HBW-Impedance-WalkTime`,
                              `Retail-Purpose-HBW` = `Retail-Purpose-HBW-Impedance-WalkTime`,
                              `Service-Purpose-HBW` = `Service-Purpose-HBW-Impedance-WalkTime`,
                              `Basic-Purpose-HBW` = `Basic-Purpose-HBW-Impedance-WalkTime`,
                              `TotEnroll-Purpose-HBW` = `TotEnroll-Purpose-HBW-Impedance-WalkTime`)
zones_b <- zones_b %>% rename(`Total Emp-Purpose-HBW` = `Total Emp-Purpose-HBW-Impedance-BikeTime`,
                              `Retail-Purpose-HBW` = `Retail-Purpose-HBW-Impedance-BikeTime`,
                              `Service-Purpose-HBW` = `Service-Purpose-HBW-Impedance-BikeTime`,
                              `Basic-Purpose-HBW` = `Basic-Purpose-HBW-Impedance-BikeTime`,
                              `TotEnroll-Purpose-HBW` = `TotEnroll-Purpose-HBW-Impedance-BikeTime`)

z_cols = c("INWINDOW", "INFOCUS", "TAZ", "Total Emp-Purpose-HBW",
           "Retail-Purpose-HBW", "Service-Purpose-HBW", "TotEnroll-Purpose-HBW",
           "geometry")
combo_w <- zones_w %>% 
  select(z_cols) %>% 
  bind_rows(
    blocks_w %>% select(c('block_id', z_cols)
                        )
    )
combo_b <- zones_b %>% 
  select(z_cols) %>% 
  bind_rows(
    blocks_b %>% select(c('block_id', z_cols)
                        )
    )

# Coerce to sf
zones_a <- st_sf(zones_a)
zones_t <- st_sf(zones_t)
combo_w <- st_sf(combo_w)
combo_b <- st_sf(combo_b)

# Map
a <- tm_shape(zones_a) + tm_polygons(col = "Total Emp-Purpose-HBW",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Access to jobs by car")

t <- tm_shape(zones_t) + tm_polygons(col = "Total Emp-Purpose-HBW",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Access to jobs by transit")

w <- tm_shape(combo_w) + tm_polygons(col = "Total Emp-Purpose-HBW",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Access to jobs by walking")

b <- tm_shape(combo_b) + tm_polygons(col = "Total Emp-Purpose-HBW",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6, 
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Access to jobs by biking")
tmap_arrange(a, t, w, b, ncol=2, nrow=2, sync=TRUE)
```


### Trip generation
Trip generation focuses on the number of trips produced by and attracted to
each zone. Trip generation estimates vary by household characteristics (size,
income, number of workers, and vehicle ownership), job types (basic sector, 
retail, service), and campus-based activities (group quarters populations and
enrollment). Person trip estimates interact with mode choice estimates to
approximate trips by mode to and from each zone prior to trip distribution.

The maps below show person trip productions (left) and attractions (right) for
the home-based work (upper) and home-based other (lower) trip purposes.

```{r, eval = TRUE}
tmap_mode("plot")
setwd(data_path)
# Read tables
zdata <- read_csv("./FinalData/lu/Base/trips_by_zone.csv")
prods = zdata %>% filter(End == "P")
attrs = zdata %>% filter(End == "A")
hbw_p = prods %>%  filter(Purpose == "HBW")
hbo_p = prods %>%  filter(Purpose == "HBO")
hbw_a = attrs %>%  filter(Purpose == "HBW")
hbo_a = attrs %>%  filter(Purpose == "HBO")

# Join data
zones_a <- left_join(
  hbw_p, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_t <- left_join(
  hbw_a, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_w <- left_join(
  hbo_p, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_b <- left_join(
  hbo_a, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))

# Coerce to sf
zones_a <- st_sf(zones_a)
zones_t <- st_sf(zones_t)
zones_w <- st_sf(zones_w)
zones_b <- st_sf(zones_b)

# Bounding box for base map
basemap_bbox <- st_bbox(
  st_transform(zones_a, 26919) %>% 
  st_buffer(41000.0) %>% 
  st_transform(4326)) 
basemap <- osm.raster(basemap_bbox, crop=TRUE, type = "cartolight", 
                      quiet = TRUE, progress = "none")

# Map
a <- tm_shape(basemap) + tm_rgb() + tm_shape(zones_a, is.master = TRUE) + 
        tm_fill(col = "Trips",
                    palette = rpg_color_pal("rpg_rainbow")(9),
                    n = 9,
                    alpha=0.4,
                    style = "pretty",
                    title = "HBW Productions")

t <- tm_shape(basemap) + tm_rgb() + tm_shape(zones_t, is.master = TRUE) + 
        tm_fill(col = "Trips",
                    palette = rpg_color_pal("rpg_rainbow")(9),
                    n = 9,
                    alpha=0.4,
                    style = "pretty",
                    title = "HBW Attractions")

w <- tm_shape(basemap) + tm_rgb() + tm_shape(zones_w, is.master = TRUE) + 
        tm_fill(col = "Trips",
                    palette = rpg_color_pal("rpg_rainbow")(9),
                    n = 9,
                    alpha=0.4,
                    style = "pretty",
                    title = "HBO Productions")

b <- tm_shape(basemap) + tm_rgb() + tm_shape(zones_b, is.master = TRUE) + 
        tm_fill(col = "Trips",
                    palette = rpg_color_pal("rpg_rainbow")(9),
                    n = 9,
                    alpha=0.4,
                    style = "pretty",
                    title = "HBO Attractions")

tmap_arrange(a, t, w, b, ncol=2, nrow=2, sync=TRUE)
```


### Mode choice
The mode choice phase estimates the share of trip productions and attractions
made by each of the six travel modes analyzed. The person trips estimated in the
trip generation phase are multiplied by each zone's mode share estimated to
determine the number of trips by mode starting and ending in each zone. 

The maps below show mode share estimates for home-based work trip productions 
for auto (driver + auto passenger), transit (either access mode), walking, and 
biking.

```{r, eval = TRUE}
tmap_mode("view")
setwd(data_path)
# Read tables
zdata <- read_csv("./FinalData/scen/Base/trips_by_mode_predist.csv")
hbw <- zdata %>% filter(Purpose == "HBW" & End == "P")
hbw_tot <- hbw %>% group_by(TAZ) %>%  summarise(all_trips = sum(trips))

auto = hbw %>%  filter(Mode == "driver" | Mode == "passenger") %>% 
  group_by(TAZ) %>% summarise(trips = sum(trips))
tran = hbw %>%  filter(Mode == "WAT" | Mode == "DAT") %>% 
  group_by(TAZ) %>% summarise(trips = sum(trips))
walk = hbw %>%  filter(Mode == "walk")
bike = hbw %>%  filter(Mode == "bike")

# Calc share
auto <- left_join(auto, hbw_tot) %>% mutate(share=trips/all_trips)
tran <- left_join(tran, hbw_tot) %>% mutate(share=trips/all_trips)
walk <- left_join(walk, hbw_tot) %>% mutate(share=trips/all_trips)
bike <- left_join(bike, hbw_tot) %>% mutate(share=trips/all_trips)

# Join data
zones_a <- left_join(
  auto, zones, c("TAZ" = "ID"))
zones_t <- left_join(
  tran, zones, c("TAZ" = "ID"))
zones_w <- left_join(
  walk, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))
zones_b <- left_join(
  bike, zones, c("TAZ" = "ID", "INWINDOW"="INWINDOW", "INFOCUS"="INFOCUS"))

# Coerce to sf
zones_a <- st_sf(zones_a)
zones_t <- st_sf(zones_t)
zones_w <- st_sf(zones_w)
zones_b <- st_sf(zones_b)

# Map
a <- tm_shape(zones_a) + tm_polygons(col = "share",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Auto share")

t <- tm_shape(zones_t) + tm_polygons(col = "share",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Transit share")

w <- tm_shape(zones_w) + tm_polygons(col = "share",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6,
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Walk share")

b <- tm_shape(zones_b) + tm_polygons(col = "share",
                                     palette = rpg_color_pal("rpg_rainbow")(9),
                                     n = 9,
                                     alpha=0.6, 
                                     style = "pretty",
                                     lwd = 0.1,
                                     title = "Bike share")
tmap_arrange(a, t, w, b, ncol=2, nrow=2, sync=TRUE)
```

### Distribution
With trip productions and attractions in each zone broken down by mode, the
distribution step focuses on estimating which OD pairs those trips travel 
between. Trips by mode are distributed to approximate typical trip lengths and
costs, but are constrained to ensure total person trip productions and
attractions match the trip generation results. 

The map below displays the distribution of trips by mode for home-based work
trips in the morning peak period.


```{r plot trips}
# Read skims
setwd(data_path)
skims <- read_csv("./FinalData/scen/Base/trips_by_mode_od_gt1.csv")
# Set lat/lon coords for trip ends
zones_centroid_latlong <- st_sf(zones %>% select(ID, geometry)) %>% 
  st_centroid(zones) %>%
  st_transform(4326) %>% 
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2]) %>% 
  st_drop_geometry()

trips_od <- left_join(skims, zones_centroid_latlong, by = c("TAZ_From" = "ID")) %>% 
  rename(f_lat = lat, f_lon = lon)

trips_od <- left_join(trips_od, zones_centroid_latlong, by = c("TAZ_To" = "ID")) %>% 
  rename(t_lat = lat, t_lon = lon) %>% 
  drop_na()
# Set line widths
breaks_trips <- classIntervals(trips_od$trips, n=7, style='fisher')
trips_od$trips_cat <- cut(trips_od$trips, breaks = breaks_trips$brks)
trips_od$width <- cut(trips_od$trips, breaks = breaks_trips$brks, labels = c(".5", "1", "2", "5", "10", "25", "40"))
trips_od$width <- as.numeric(as.character(trips_od$width))
trips_od <- drop_na(trips_od)
# Make map
mapdeck(token = "pk.eyJ1IjoiamxlaG1hbjkzIiwiYSI6ImNrYjltdGtkYzAybmMyc25xbGV4ODBjeTYifQ.WjcnEpbpl4mKAjHRb8JD0w", 
        style = mapdeck_style('dark'), pitch = 90) %>%
  add_arc(
    data = trips_od,
    origin = c('f_lon','f_lat'),
    destination = c('t_lon','t_lat'),
    stroke_from = 'Mode',
    stroke_to = 'Mode',
    stroke_width = 'width', 
    legend = TRUE,
    digits = 4,
    brush_radius = 1000,
    palette = "matlab_like2"
  )

```


### TNC Trips
The final step in the accessibility-based travel model focuses on trips made
using transportation network companies (TNC's) like Uber or Lyft. This is a
post processing step that estimates which trips are likely to switch to TNC
travel based on assumed TNC costs, trip purpose, and the original mode selected.

The maps below show TNC trip productions (left) and attractions (right) as
total estimated TNC trips (upper) and the TNC share of total trips (lower).

```{r, eval = TRUE}
tmap_mode("plot")
setwd(data_path)
# Read tables
zdata <- read_csv("./FinalData/scen/Base/TNC_zone_summary_AM.csv")
hbw <- zdata %>% filter(Purpose == "HBW" & INWINDOW == 1)

prods = hbw %>% 
  filter(FT == F) %>% 
  group_by(TAZ) %>% 
  summarise(tnc_trips = sum(tnc_trips), total_trips = sum(total_trips)) %>% 
  mutate(pct = tnc_trips/total_trips)

attrs = hbw %>% 
  filter(FT == T) %>% 
  group_by(TAZ) %>% 
  summarise(tnc_trips = sum(tnc_trips), total_trips = sum(total_trips)) %>% 
  mutate(pct = tnc_trips/total_trips)

# Join data
prods <- left_join(
  prods, zones, c("TAZ" = "ID"))
attrs <- left_join(
  attrs, zones, c("TAZ" = "ID"))

# Coerce to sf
prods <- st_sf(prods)
attrs <- st_sf(attrs)

# Bounding box for base map
basemap_bbox <- st_bbox(
  st_transform(prods, 26919) %>% 
  st_buffer(1400.0) %>% 
  st_transform(4326)) 
basemap <- osm.raster(basemap_bbox, crop=TRUE, type = "cartolight", 
                      quiet = TRUE, progress = "none")

# Map
a <- tm_shape(basemap) + tm_rgb() + tm_shape(prods, is.master = TRUE) + 
        tm_polygons(col = "tnc_trips",
                palette = rpg_color_pal("rpg_rainbow")(9),
                n = 9,
                alpha=0.4,
                style = "pretty",
                title = "TNC Productions")

t <- tm_shape(basemap) + tm_rgb() + tm_shape(attrs, is.master = TRUE) + 
        tm_polygons(col = "tnc_trips",
                palette = rpg_color_pal("rpg_rainbow")(9),
                n = 9,
                alpha=0.4,
                style = "pretty",
                title = "TNC Attractions")

w <- tm_shape(basemap) + tm_rgb() + tm_shape(prods, is.master = TRUE) + 
        tm_polygons(col = "pct",
                palette = rpg_color_pal("rpg_rainbow")(9),
                n = 9,
                alpha=0.4,
                style = "pretty",
                title = "% TNC Productions")

b <- tm_shape(basemap) + tm_rgb() + tm_shape(attrs, is.master = TRUE) + 
        tm_polygons(col = "pct",
                palette = rpg_color_pal("rpg_rainbow")(9),
                n = 9,
                alpha=0.4,
                style = "pretty",
                title = "% TNC Attractions")

tmap_arrange(a, t, w, b, ncol=2, nrow=2, sync=TRUE)
```




